Probe or Wait: Handling tail losses using Multipath TCP

### 2017-03-09 main.tex
Update after Per comments

The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics. Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths. Quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics. MPTCP standard suggests a more conservative approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.

@@

In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values ranging from 0 to 4 with 3 being the Linux default. The default setting value enables both
ER and TLP. TCP in other operating systems might not use ER and TLP. We use RTO to refer to a sysctl setting of 0 and TLP to refer to a sysctl setting of 3 throughout the paper.
In Linux, disabling ER and TLP gives the performance of RTO. In Linux default TCP, enabling delayed ER gives the combined performance of RTO, delayed ER and TLP.
In RTO, probe loss scenario is equivalent to losing the retransmitted packet.

The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}. These scenarios are chosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays. For each experiment trail, we send two bursts of data from server to client, each burst containing 80 packets. First burst allows the MPTCP connection to be setup taking in to account that second flow is added after some time and packets. We calculate the burst completion time for second burst that has specified tail losses.

@@

Experiments performed on CORE emulator with the setup mentioned in section~\ref{exsetup}. MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler.

The burst completion times on single packet tail loss calculated for each delay configuration and retransmission 
setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.

>>>>

The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics. Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always to follow TCP semantics. MPTCP standard suggests a more conservative approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.

@@

In Linux, TCP supports multiple retransmission mechanisms that can be controlled by a system setting. The default setting in Linux enables both
ER and TLP, which may not be the case with other operating systems. In this paper, we use TLP to denote the default Linux setting and RTO to denote the use of retransmission timeout only.

The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher aysmmetry in the link delays. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}. For each experiment trail, we send two bursts of data from server to client, each burst containing 80 packets. First burst allows the MPTCP connection to be setup taking in to account that second flow is added after some time and packets. We calculate the burst completion time for second burst that has specified tail losses.

@@

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP.
